---
title: "Colon Cancer classification"
author: "David Cáceres"
date: "2023-05-09"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(reshape2)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggsci)
library(patchwork)
library(cowplot)
```


# Reading data
```{r}
quality<-read.table(file="Quality features/Quality_table.tsv", header = TRUE, sep = "\t")
quality$Batch<-as.factor(quality$Batch)
rownames(quality)<-quality$Sample
quality<-quality[,-1]
```

# Mapping distribution

## Dataframe format

```{r}
mapping<-data.frame(rownames(quality), quality$Batch, quality$Uniquely.mapped.reads.perc, quality$perc.of.reads.mapped.to.multiple.loci,quality$perc.of.reads.mapped.to.too.many.loci, quality$perc.of.reads.unmapped..too.short, quality$perc.of.reads.unmapped..other)

colnames(mapping)<-c("sample",  "Batch", "Uniquely_mapped", "Mapped_multiple_loci","Mapped_many_loci", "Unmapped_short", "Unmaped_other")

mapped<-melt(setDT(mapping),id.vars = c("sample", "Batch"))
```

## Mapping distribution plot
```{r}
mapped <- mapped |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Uniquely_mapped", 0, value), FUN = sum))


p1 <- ggplot(mapped,aes(x = sample1, y = value, fill = factor(variable, levels=c("Unmaped_other", "Unmapped_short","Mapped_many_loci", "Mapped_multiple_loci", "Uniquely_mapped")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
    labs(x = NULL, y= "% Reads" , title="Mapped reads% Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_locuszoom(labels=c("Unmaped other","Unmaped short","Mapped many loci", "Mapped multiple", "Uniquely mapped"
                                     ))
mapped1 <- mapped |>
    distinct(sample, sample1, Batch)

p_axis <- ggplot(mapped1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p1_q<-p1 / p_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p1_q
```


# Trimming distribution

## Dataframe format

```{r}
Adapter_found<-data.frame(rownames(quality),quality$Batch, quality$Adapter.found.perc, quality$Adapter.not.found)

colnames(Adapter_found)<-c("sample", "Batch", "Adapter_found_perc","Adapter_not_found_perc")

Adapter_found<-melt(setDT(Adapter_found),id.vars = c("sample", "Batch"))
```

## Trimming distribution plot

```{r}
Adapter_found <- Adapter_found |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Adapter_found_perc", 0, value), FUN = sum))


p2 <- ggplot(Adapter_found,aes(x = sample1, y = value, fill = factor(variable, levels=c("Adapter_not_found_perc", "Adapter_found_perc")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
    labs(x = NULL, y= "% Reads" , title="Adapter Reads% Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_uchicago(labels=c("Adapter not found", "Adapter found"))

Adapter_found1 <- Adapter_found |>
    distinct(sample, sample1, Batch)

p2_axis <- ggplot(Adapter_found1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p2_q<-p2 / p2_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p2_q
```

# Trimming lexogen adapter
## Dataframe format

```{r}
AdapterLex_found<-data.frame(rownames(quality),quality$Batch, quality$Reads.main.adapter.perc, quality$Other.adapter.perc)

colnames(AdapterLex_found)<-c("sample", "Batch", "AdapterLex_found_perc","Other_Adapter_found_perc")

AdapterLex_found<-melt(setDT(AdapterLex_found),id.vars = c("sample", "Batch"))
```

## Lexogen adapter distribution plot

```{r}
AdapterLex_found <- AdapterLex_found |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "AdapterLex_found_perc", 0, value), FUN = sum))


p3 <- ggplot(AdapterLex_found,aes(x = sample1, y = value, fill = factor(variable, levels=c("Other_Adapter_found_perc", "AdapterLex_found_perc")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0), breaks = seq(0, 100, by = 10)) +
    labs(x = NULL, y= "% Reads" , title="Lexogen Adapter Reads% Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_uchicago(labels=c("Other adapter found", "Lexogen adapter found"))

AdapterLex_found1 <- AdapterLex_found |>
    distinct(sample, sample1, Batch)

p3_axis <- ggplot(AdapterLex_found1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p3_q<-p3 / p3_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p3_q
```


# Total Reads distribution
## Dataframe format

```{r}
Total_reads<-data.frame(rownames(quality),quality$Batch, quality$Total.input.reads)

colnames(Total_reads)<-c("sample", "Batch", "Total_reads")

Total_reads<-melt(setDT(Total_reads),id.vars = c("sample", "Batch"))
```


## Total reads distribution plot

```{r}
Total_reads <- Total_reads |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Total_reads", 0, value), FUN = sum))


p4 <- ggplot(Total_reads,aes(x = sample1, y = value, fill = factor(variable, levels=c("Total_reads")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = NULL, y= "Reads" , title="Total Imput Reads Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_d3(labels=c("Total Reads"))

Total_reads1 <- Total_reads |>
    distinct(sample, sample1, Batch)

p4_axis <- ggplot(Total_reads1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p4_q<-p4 / p4_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p4_q
```




# Total Reads after adapter trimming
## Dataframe format

```{r}
Total_reads_trim<-data.frame(rownames(quality),quality$Batch, quality$FastQC_mqc.generalstats.fastqc.total_sequences)

colnames(Total_reads_trim)<-c("sample", "Batch", "Total_sequences")

Total_reads_trim<-melt(setDT(Total_reads_trim),id.vars = c("sample", "Batch"))
```


## Total reads after trimming distribution plot

```{r}
Total_reads_trim <- Total_reads_trim |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Total_sequences", 0, value), FUN = sum))


p5 <- ggplot(Total_reads_trim,aes(x = sample1, y = value, fill = factor(variable, levels=c("Total_sequences")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = NULL, y= "Reads" , title="Total Reads distribution after trimming", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_d3(labels=c("Total Reads"))

Total_reads1 <- Total_reads |>
    distinct(sample, sample1, Batch)

p5_axis <- ggplot(Total_reads1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p5_q<-p5 / p5_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p5_q
```


# Duplicate reads distribution
## Dataframe format

```{r}
Duplicate<-data.frame(rownames(quality),quality$Batch, quality$FastQC_mqc.generalstats.fastqc.percent_duplicates)

colnames(Duplicate)<-c("sample", "Batch", "Duplicates")

Duplicate<-melt(setDT(Duplicate),id.vars = c("sample", "Batch"))
```



```{r}
Duplicate <- Duplicate |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Duplicates", 0, value), FUN = sum))


p6 <- ggplot(Duplicate,aes(x = sample1, y = value, fill = factor(variable, levels=c("Duplicates")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = NULL, y= "Reads%" , title="Duplicates% Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_d3(labels=c("Reads%"))

Duplicate1 <- Duplicate |>
    distinct(sample, sample1, Batch)

p6_axis <- ggplot(Duplicate, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p6_q<-p6 / p6_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p6_q
```

# Read length distribution

## Dataframe format

```{r}
Length<-data.frame(rownames(quality),quality$Batch, quality$FastQC_mqc.generalstats.fastqc.avg_sequence_length)

colnames(Length)<-c("sample", "Batch", "Length")

Length<-melt(setDT(Length),id.vars = c("sample", "Batch"))
```


```{r}
Length <- Length |> 
  mutate(sample1 = reorder(sample, -ifelse(!variable %in% "Length", 0, value), FUN = sum))


p7 <- ggplot(Length,aes(x = sample1, y = value, fill = factor(variable, levels=c("Length")), width=1.05)) +
    geom_bar( stat = "identity", position = position_stack()) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(x = NULL, y= "Reads length (bp)" , title="Read length Distribution", 
         fill="Quality Control") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.length.x = unit(0, "pt")) +
    theme(
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent'),
        plot.title = element_text(face="bold")
    ) + scale_fill_d3(labels=c("Read length"))

Length1 <- Length |>
    distinct(sample, sample1, Batch)

p7_axis <- ggplot(Length1, aes(x = sample1, y = factor(1), 
                                 fill = Batch)) +
    geom_tile(width = 1) +
    theme_void() +
    theme(axis.title.x = element_text()) + theme(legend.position = "right")  +
    labs(x="Batch Annotation", fill="Batch") 

p7_q<-p7 / p7_axis + plot_layout(heights = c(8,1))  + scale_fill_jama()
p7_q
```

```{r}
plot_grid(p1_q,p4_q, p5_q, p6_q, p2_q, p3_q)
```

